---
title: "Enterprise Surveys and Expopriation Risk"
author: "Robert Kubinec"
date: "December 18, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
require(dplyr)
require(tidyr)
require(ggplot2)
require(haven)
require(lubridate)
```

## Data

I am using the full release of World Bank Enterprise Survey data:

```{r data_load}
wb_data <- readRDS('data/wb_data.rds') %>% as_factor
# get question labels
question_labels <- sapply(wb_data,attr,'label')
all_qs <- data_frame(q_num=question_labels,q_label=names(question_labels))
```

It contains `r nrow(wb_data)` observations from `r length(unique(wb_data$country))` countries.

Unfortunately, the data do not contain questions about expropriation risk faced by firms. I go over some proxies that come up in this data that we can use to get an initial sense of things:

## Indicators: Corruption 

Some of the following indicators reflect firm perceptions of corruption:


----------------------------------------------------------------------
Number  Question                     Answers (in increasing order)
------  ---------------------------  ----------------------------------
c5      Informal Gift/Payment         No (66%), Yes (12.4%)
        Expected Or Requested For 
        An Electrical Connection?
 
c14     When You Applied For A        No (52%), Yes (8.7%)
        Water Connection, Was 
        An Informal Gift/Payment 
        Requested?

c21     When You Applied For A        No (72%), Yes (5.3%)
        Telephone Connection 
        Was An Informal 
        Gift Requested?

d5a     In Clearing Exports           No (77%), Yes (11%)
        Through Customs,Informal 
        Gift/Payment Expected 
        Or Requested?
        
d15a    In Clearing Imports           No (86%), Yes (9.5%)
        Through Customs,
        Informal Gift/Payment 
        Expected Or Requested?
        
g4      Informal Gift/Payment         No (60%), Yes (15%)
        Expected Or Requested 
        For A Construction-
        Related Permit

j5      In Any Of These               No (80%), Yes (13%)
        Inspections Was A 
        Gift/Informal 
        Payment Requested ?

j6      % Of Contract Value           Mean = 4.7%, 
        Avg. Firm Pays In             Median = 5%,  
        Informal Gifts To             SD=0.6%
        Govt To Secure 
        Contract?

j7a     Percent Of Total              Mean=4.5%,     
        Annual Sales Paid             Median=5%,
        In Informal                   SD=4%
        Payments
        
j7b     Total Annual                  Mean=6.9%,
        Informal Payment              Median=6%,
                                      SD=11%

j12     When You Applied              No (66%), Yes (9.8%)
        For An Import                 
        License, 
        Was An Informal 
        Gift/Payment 
        Requested?

j15     When You                      No (73%), Yes (14%)
        Applied For 
        Operating License 
        Was An Informal 
        Gift/Payment 
        Requested?

j30f    How Much Of                   No obstacle (30%),
        An Obstacle:                  Major obstacle (18%),
        Corruption                    Very severe obstacle (17%),
                                      Moderate obstacle (16%),
                                      Minor obstacle (15%)

m1a     Biggest Obstacle              Corruption (8%),
        Affecting The                 Largest = Access to     
        Operation Of                  Finance (15%)  
        This Establishment
                                      
--------------------------------------------------------------------------

## Indicators: Court System


----------------------------------------------------------------------
Number  Question                     Answers (in increasing order)
------  ---------------------------  ----------------------------------
h30     How Much Of An Obstacle:      No obstacle (48%), 
        Courts                        Minor obstacle (18%),
                                      Moderate obstacle (13%),
                                      Major obstacle (9%),
                                      Very severe obstacle (5%)
                                      
j1a     The court system is fair,     Tend to Agree (31%), 
        impartial and uncorrupted     Tend to Disagree (29%),
                                      Strongly Disagree (23%)
                                      Strongly Agree (12%)
                                      
h7a     Government officials’         Tend to Agree (28%),
        interpretations of the        Tend to Disagree (27%),
        laws and regulations          Strongly Disagree (26%),
        affecting this                Strongly Agree (11%)
        establishment are 
        consistent and 
        predictable.
-------------------------------------------------------------------------

## Indicators: Crime & Theft

----------------------------------------------------------------------
Number  Question                     Answers (in increasing order)
------  ---------------------------  ----------------------------------
i1      In Last Fy, Did               Yes (62%), No (38%)
        This Establishment Pay 
        For Security?

i4a     Losses Due To Theft,          Mean=4.5%,
        Robbery, Vandalism Or         Median=4%,
        Arson In Last Fiscal          SD=9.2%
        Yr (% Of Sales)

i30     How Much Of An                No obstacle (41%),
        Obstacle: Crime, Theft        Minor obstacle (23%),
        And Disorder?                 Moderate Obstacle (15%)
                                      Major obstacle (12%),
                                      Very severe obstacle (8%)
-------------------------------------------------------------------------


## Indicators: Government & Politics

----------------------------------------------------------------------
Number  Question                     Answers (in increasing order)
------  ---------------------------  ----------------------------------
j30e    How Much Of An Obstacle:      No obstacle (33%),
        Political Instability         Major obstacle (18%),
                                      Moderate Obstacle (17%),
                                      Minor Obstacle (15%),
                                      Very severe obstacle (14%)
                                      
j30c    How Much Of An Obstacle:      No obstacle (43%),
        Business Licensing And        Minor obstacle (21%),  
        Permits                       Moderate obstacle (18%),
                                      Major obstacle (10%),
                                      Very severe obstacle (4%)
                                    
j2      What % Of Senior              Mean=34%,
        Management Time Was           Median=18%,
        Spent In Dealing With         SD=32%
        Govt Regulations?
                                    
-------------------------------------------------------------------------
                                      
--------------------------------------------------------------------------

## Run Some IRT Analysis

Now we can use `mirt` to scale the countries (groups) using the factors above.

```{r to_mirt}
require(stringr)
require(forcats)
require(mirt)
require(missRanger)

recode_func <- function(col) {
  levels(col)[grepl(pattern = 'Spontaneous|-[0-9]',
                      x=levels(col))] <- NA
  return(col)
}
just_mirt <- select(wb_data,
                    country,
                    j30e,
                    j2,
                    j30c,
                    i30,
                    i4a,
                    j7a,
                    i1,
                    h30,
                    j1a,
                    h7a,
                    m1a,
                    j30f,
                    j15,
                    j12,
                    j5,
                    j6,
                    g4,
                    d15a,
                    d5a,
                    c21,
                    c14,
                    c5) %>% 
  mutate(year=str_extract(country,'[0-9]{4}'),
         country=str_replace(country,'[0-9]{4}|-',
                             "")) %>% 
  mutate_all(funs(recode_func))

group_by(just_mirt,country) %>% 
  summarize(mean_time=mean(as.numeric(j2),na.rm = T)) %>% 
  arrange(mean_time) %>% 
  knitr::kable(.,caption = "Time Spent Dealing with Regulations")

group_by(just_mirt,country,j30e) %>% 
  count %>% 
  filter(!is.na(j30e)) %>% 
    group_by(country) %>% 
  mutate(n=n/sum(n)) %>% 
  arrange(j30e,desc(n)) %>% 
  knitr::kable(.,caption='How Much of an Obstacle Political Instability')
group_by(just_mirt,country,j30c) %>% 
  count %>% 
    filter(!is.na(j30c)) %>% 
    group_by(country) %>% 
  mutate(n=n/sum(n)) %>% 
  arrange(j30c,desc(n)) %>% 
  knitr::kable(.,caption='How Much of an Obstacle Business Licenses')
group_by(just_mirt,country,i30) %>% 
  count %>% 
    filter(!is.na(i30)) %>% 
    group_by(country) %>% 
  mutate(n=n/sum(n)) %>% 
  arrange(i30,desc(n)) %>% 
  knitr::kable(.,caption='How Much of an Obstacle Crime')
group_by(just_mirt,country,i1) %>% 
  count %>% 
    filter(!is.na(i1)) %>% 
  group_by(country) %>% 
  mutate(n=n/sum(n)) %>% 
  arrange(i1,desc(n)) %>% 
  knitr::kable(.,caption='Did Establishment pay for security?')
group_by(just_mirt,country,j30f) %>% 
  count %>% 
    filter(!is.na(j30f)) %>% 
  group_by(country) %>% 
  mutate(n=n/sum(n)) %>% 
  arrange(j30f,desc(n)) %>% 
  knitr::kable(.,caption='Corruption An Obstacle')
group_by(just_mirt,country,h30) %>% 
  count %>% 
    filter(!is.na(h30)) %>% 
  group_by(country) %>% 
  mutate(n=n/sum(n)) %>% 
  arrange(h30,desc(n)) %>% 
  knitr::kable(.,caption='Courts An Obstacle')
```


